Catching fire: Undefined behaviour for equational reasoning of lazy languages with strictness annotations

### 2021-01-21 paper/paper.tex
add PDF
add factorial worker/wrapper example that becomes a constant value

@@

>>>>

The transformation of floating out common patterns works equally well
for branches as it does for sum types, since a case analysis on a sum
type is the same as a branch on the tag.

@@

\subsection{A toy example: eliminating wrapper overhead}

\begin{lstlisting}[language=llvm,caption="Original program with wrapper and laziness overhead"]
module {
// fact 0 = 1
// fact n = n*fact(n-1)
func @f (%i : !lz.thunk<i64>) -> i64 {
  %ival = lz.force(%i):i64
  %out = lz.caseint %ival 
   [0 ->   { 
    ^entry:
    %c1 = constant 1 : i64
    lz.return %c1 : i64
   }]
   [@default -> {
    ^entry:
    %c1 = constant 1 : i64
    %idec = subi %ival, %c1 : i64
    %idecthnk = lz.thunkify(%idec : i64) 
    %f = constant @f : (!lz.thunk<i64>) -> i64
    %f_idec_thnk = lz.ap(%f: (!lz.thunk<i64>) -> i64, 
        %idecthnk)
    %f_idec_val = lz.force(%f_idec_thnk) : i64
    %prod = muli %ival, %f_idec_val : i64
    lz.return %prod : i64
   }]
  return %out : i64
}

func @c8 () -> i64 {
  %v = std.constant 8 : i64
  return %v: i64
}

func @main() -> i64 {
  %f = constant @f : (!lz.thunk<i64>) -> i64
  %c8f = constant @c8 : () -> i64
  %c8t = lz.ap(%c8f : () -> i64)
  %outt = lz.ap(%f: (!lz.thunk<i64>) -> i64, %c8t)
  %out = lz.force(%outt) : i64
  return %out : i64
}
}

\end{lstlisting}

After worker/wrapper, we get the optimized program:

\begin{lstlisting}[language=llvm,caption="factorial that has been worker/wrapper'd"]
module  {
  func @c8() -> i64 {
    %c8_i64 = constant 8 : i64
    return %c8_i64 : i64
  }
  func @main() -> i64 {
    %c7_i64 = constant 7 : i64
    %f = constant @frec_force_outline : (i64) -> i64
    %c8_i64 = constant 8 : i64
    %0 = "lz.apEager"(%f, %c7_i64) : ((i64) -> i64, i64) -> i64
    %1 = muli %0, %c8_i64 : i64
    return %1 : i64
  }
  func @frec_force_outline(%arg0: i64) -> i64 {
    %c1_i64 = constant 1 : i64
    %f = constant @frec_force_outline : (i64) -> i64
    %0 = "lz.caseint"(%arg0) ( {
      lz.return %c1_i64 : i64
    },  {
      %1 = subi %arg0, %c1_i64 : i64
      %2 = "lz.apEager"(%f, %1) : ((i64) -> i64, i64) -> i64
      %3 = muli %arg0, %2 : i64
      lz.return %3 : i64
    }) {alt0 = 0 : i64, alt1 = @default} : (i64) -> i64
    return %0 : i64
  }
\end{lstlisting}

\sid{Explain how this happens}
We will explain this transformation step by step. Finally, when lowered to
LLVM, LLVM's strong loop analyses is able to compute a closed form for 
\texttt{frec\_force\_outline}, ending in a constant time computation.

\subsection{A toy example: eliminating laziness and exposing loop optimisation}
